---
title: "ggplot2 - Bar Charts"
author: "Vincent La"
date: "May 25, 2017"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```

# Bar Charts
This is a short introduction and reference guide to bar charts in ggplot2 which assumes some experience in R but very little using ggplot2.

## Visualizing Sasquatch Sightings
The dataset we will be working with is fictious, just like Sasquatch. Suppose we had a table of months and the number of Sasquatch sightings per month.

```{r}
sas_dat <- data.frame(
  months = c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep",
             "Oct", "Nov", "Dec"),
  sightings = c(10, 20, 15, 25, 30, 54, 23, 123, 12, 0, 11, 12)
)

sas_dat
```

## Using pre-calculated counts with `stat='identity'`
The `stat='identity'` option tells ggplot2 that the y-values should be left as is. On the other hand, ggplot's default behavior is to assume that you are giving it data that needs to be counted.

```{r}
sas_plot <- ggplot(data=sas_dat, aes(x=months, y=sightings)) +
  geom_bar(stat="identity")

sas_plot
```

### Re-Ordering
```{r}
month_names <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep",
                 "Oct", "Nov", "Dec")

sas_dat <- data.frame(
  months = factor(month_names, levels=month_names),              # Use levels argument
  sightings = c(10, 20, 15, 25, 30, 54, 23, 123, 12, 0, 11, 12)
)

ggplot(data=sas_dat, aes(x=months, y=sightings)) +
  geom_bar(stat="identity")
```

## Letting ggplot2 count for you with `stat='bin'`
```{r}
month <- c()

set.seed(420)

for (m in c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep",
             "Oct", "Nov", "Dec")) {
  month <- append(month, rep(m, runif(1, min=0, max=20)))
}
```


```{r}
dat <- data.frame(
 month = month
)

my_plot <- ggplot(data=dat, aes(x=month)) + geom_bar()
my_plot
```

## Fancying Up our Plots
### Renaming the X and Y Axes
Because ggplot2 names the axes after our data frame's column names, directly changing them will also change the axis labels. Although this is a rather inefficient way of renaming the columns, it is included because it demonstrates use of the sometimes useful built-in `names()` function.

```{r}
sas_dat3 <- sas_dat

# Use names() function to directly access column names
names(sas_dat3)
```

```{r}
# Change column names
names(sas_dat3) = c("Month", "Sasquatch Sightings")
```

However, this method can be inconvenient as you will also have to change how you refer to your columns as well. Notice that the argument y=`Sasquatch Sightings` was used to escape the scape in the column name.

```{r}
my_plot <- ggplot(data=sas_dat3, aes(x=Month, y=`Sasquatch Sightings`)) +
  geom_bar(stat="identity")

my_plot
```

### Renaming the X and Y Axes: The Easier Way
This method only adds an extra line of code (if you use the same whitespace rules as I do).

```{r}
ggplot(data=sas_dat, aes(x=months, y=sightings)) +
  geom_bar(stat="identity") +
  xlab("Months") + ylab("Sasquatch Sightings")      # Change x and y labels
```